---
title: "Week 1 Challenge Instructions"
author: "Oriol Vallès Codina"
description: "Loading Data and Creating a Network"
date: "02/13/2023"
format:
  html:
    toc: true
    code-fold: true
    code-copy: true
    code-tools: true
# editor: visual
categories: 
  - challenge_1
  - instructions
  # - railroads
  # - faostat
  # - wildbirds
---

```{r}
#| label: setup
#| include: false
#| 
```

## Challenge Overview

Today's challenge is to

1) read in a dataset, and

2) create a network object

## Load the Data

Read in one (or more) of the following data sets, using the correct R package and command.

-   got_marriages.csv
-   fish_encounters dataset (available in the `tidyr` package)
-   got_like_dislike.csv

Find the `_data` folder, located inside the `posts` folder. Then you can read in the data, using base `read.csv` or `read_csv` standard tidy read command to read Comma Separated Values files or, alternatively, read the data in directly from loading a package.

```{r}
```

Add any comments or documentation as needed. More challenging data sets may require additional code chunks and documentation.

## Create a Network

Load the package `igraph` and create an `igraph` object (i.e. a graph or network) in the form of an edge list. The command may vary whether the data is given as a list of connections or an adjacency matrix. Is the network directed or undirected; weighted or unweighted; unimodal or bipartite? Can you plot it?

```{r}
#| label: summary
```