---
title: "Challenge 1 Kruzlic"
author: "Bryn Kruzlic"
description: "Loading Game of Thrones marriages and Creating a Network"
date: "02/28/2023"
format:
  html:
    toc: true
    code-fold: true
    code-copy: true
    code-tools: true
# editor: visual
categories: 
  - challenge_1
  - instructions
  # - railroads
  # - faostat
  # - wildbirds
---

```{r}
#| label: setup
#| include: false
#| 
```


### Load in libraries

```{r}
# load in libraries
knitr::opts_chunk$set(echo = TRUE)
library(igraph)
```
### Load in data set. This is the Game of Thrones marriages dataset

```{r} 
# read in data set

library(readr)
got_marriages <- read_csv("_data/got/got_marriages.csv")
View(got_marriages)
```

### Create a Network

```{r} 
# plot the marriages using igraph

#| label: summary
marriages.ig <- graph_from_data_frame(got_marriages, directed = FALSE)
plot.igraph(marriages.ig, vertex.size=13)
```

#### Is the network directed or undirected? The answer is false. 
#### Is the network weighted or unweighted? The answer is false. 
#### Is the network unimodal or bipartite?  The answer is false. 

```{r}
is.bipartite(marriages.ig)
is.directed(marriages.ig)
is.weighted(marriages.ig)
```
The network is not directed, weighted, or bipartite. 


### Now, let's take a look at the vertices and edges of the network.

```{r}
vcount(marriages.ig)
ecount(marriages.ig)
```
According to the above code, there is a total of 20 vertices and 255 edges. 

### Other important factors to consider
```{r}
table(got_marriages$Type)
table(got_marriages$Generation)
table(got_marriages$From)
```

