---
title: "Week 1 Challenge"
author: "Claire Battaglia"
desription: "Loading data & creating a network"
date: "February 22, 2023"
format:
  html:
    toc: true
    code-fold: true
    code-copy: true
    code-tools: true
categories:
  - Claire Battaglia
  - challenge1

---

```{r}
#| label: setup
#| warning: false

library(tidyverse)
library(igraph)

knitr::opts_chunk$set(echo = TRUE)
```

## Data

This dataset represents the teams in Game of Thrones.

```{r data got}
# read in data
teams <- read_csv("_data/got/got_teams.csv", show_col_types = FALSE)

# specify from, to fields
teams <- teams %>%
  rename(from = team, to = joins) %>%
  relocate(from, to, alliance, gh, .before = id)

# preview
head(teams, 25)
```

Now I'll create a network object.

```{r network object got}
# create network object
teams.ig <- graph_from_data_frame(teams, directed = FALSE)

# view
teams.ig
```

This tells us a few things.

1. The network is *undirected*.
1. The network is *named*. I.e. the nodes have names.
1. The edges are not weighted.
1. It is not bipartite.
1. There are 66 nodes/vertices.
1. There are 33 edges/ties.

We can explore the nodes and edges further.

```{r nodes and edges got}
# get nodes
V(teams.ig)

# get edges
E(teams.ig)
```

Let's try to plot this.

```{r plot got}
# not sure what this does but I found it on the internet so why not :)
teams.ig <- simplify(teams.ig, remove.multiple = F, remove.loops = T)

# assign weight
E(teams.ig)$weight <- E(teams.ig)$alliance

# plot network
plot(teams.ig,
     vertex.size = 10,
     vertex.color = "maroon",
     vertex.label.cex = .5,
     main = "Game of Thrones Teams")
```

That looks... sort of useless (but vaguely like a network!).




