---
title: "Week 1 Challenge"
author: "Jyoti Rani"
description: "Loading Data and Creating a Network"
date: "02/22/2023"
format:
  html:
    toc: true
    code-fold: true
    code-copy: true
    code-tools: true
# editor: visual
categories: 
  - challenge_1
  - Loading data and creating network
  - Jyoti Rani
---

```{r}
#| label: setup
#| include: false

```

## Challenge Overview

Today's challenge is to

1) read in a dataset, and

2) create a network object

```{r}
library(tidyverse)

```


## Load the Data

Read in one (or more) of the following data sets, using the correct R package and command.

-   got_like_dislike.csv

Find the `_data` folder, located inside the `posts` folder. Then you can read in the data, using base `read.csv` or `read_csv` standard tidy read command to read Comma Separated Values files or, alternatively, read the data in directly from loading a package.

```{r}
got_like_dislike <- read_csv("_data/got/got_like_dislike.csv")
```

### The dataset has 46 rows and 49 columns

## Create a Network

Load the package `igraph` and create an `igraph` object (i.e. a graph or network) in the form of an edge list. The command may vary whether the data is given as a list of connections or an adjacency matrix. Is the network directed or undirected; weighted or unweighted; unimodal or bipartite? Can you plot it?

```{r}
library(igraph)
```
### From the below commands we know that the data is undirected, unmweighted and unimodal.

The plot is below:

```{r}
got_like_dislike.ig <- graph_from_data_frame(got_like_dislike, directed = FALSE)
is_directed(got_like_dislike.ig)
is_weighted(got_like_dislike.ig)
is_bipartite(got_like_dislike.ig)
plot.igraph(got_like_dislike.ig)
```