---
title: "Week 3 Challenge"
author: "Ben Ramsey"
description: "Degree and Density of a Network"
date: "03/06/2023"
format:
  html:
    toc: true
    code-fold: true
    code-copy: true
    code-tools: true
# editor: visual
categories: 
  - challenge_3
  - instructions
  # - railroads
  # - faostat
  # - wildbirds
---

```{r}
#| label: setup
#| include: false

library(tidyverse)
library(igraph)
library(statnet)
library(GGally)
```

## Challenge Overview

Describe the many measures of degree, as well as density, of a network and compare 

## Degree

```{r}
climate <- read.csv("_data/climate0205-allop.csv", sep =  ";") %>% subset(select = -c(X))
climate.ig <- graph_from_adjacency_matrix(as.matrix(climate), weighted = TRUE, mode = "directed", add.rownames = TRUE)
climate.net <- as.network(as.matrix(climate), directed = TRUE, weighted = TRUE)
ggnet2(climate.net, label = colnames(climate))
```

If you have not done it before, evaluate the structure of the network (number of edges and vertices, dyad and triad census, etc.).

```{r}
print(climate.net)
dyad.census(climate.net)
triad.census(climate.net, mode = "graph")
```
This is the Swiss climate policy network, which has 34 vertices and 291 edges. 65 dyads are mutual and 161 are asymmetric. There are 158 triads with 3 edges, and 787 triads with 2 edges. 

Compute the many measures of _degree_ of the network of your choice, most preferably directed. Create a data frame called `nodes` where each row corresponds to a node and each column to an attribute of the node, namely the id, name or label, and different measures of degree: total, out-degree, and in-degree. What is the average degree of the network?

```{r}
climate.nodes = data.frame(name = climate.net %v% "vertex.names", degree = degree(climate.net), indegree = degree(climate.net, cmode = "indegree"), outdegree = degree(climate.net, cmode = "outdegree"))

mean(climate.nodes$degree)
```
The average degree is 17.12. 

Compute the distributions of those measures (or histograms if your network is small). What does this tell us about the structure of the network?

```{r}
hist(climate.nodes$degree)
hist(climate.nodes$indegree)
hist(climate.nodes$outdegree)

summary(climate.nodes)
```
Most of the vertices in the network have a low degree. 

## Density

Compute the density of the network. Is this a global or local measure? Does it have a relationship with average degree?

```{r}
network.density(climate.net)
```
The density is .259, is a global measure and it does have a relationship with average degree, because it is a measure of the existing edges out of all the possible edges. 

## Random Network

Create a random (Erdos-Renyi) network with the same number of nodes and edges than the network of your choice. On `igraph`, the necessary commands are `random.graph.game(n, p.or.m, type = c("gnp", "gnm"), directed = FALSE, loops = FALSE)` (deprecated), `sample_gnp(n, p, directed = FALSE, loops = FALSE)` or `sample_gnm(n, m, directed = FALSE, loops = FALSE)`. The `p.or.m` argument is to specify the probability of an edge $p$ or the number of edges $m$.
  
Compare the densities, dyad and triad censuses, and degree distributions, with the measures of the network of your choice. Does the comparison us something about the network of your choice? 


```{r}
set.seed(1)
random.ig <- sample_gnm(34, 291)
dyad_census(random.ig)
triad_census(random.ig)
edge_density(random.ig)
random.nodes <- data.frame(degree = igraph::degree(random.ig))
hist(random.nodes$degree)
```
The Swiss climate policy network has a long right tail when it comes to total degree and it is less dense compared to the randomly generated graph.  
