---
title: "Week 1 Challenge Solutions"
author: "Oriol Vallès Codina"
editor: visual
format:
  html:
    df-print: paged
    toc: true
    code-fold: false
    code-copy: true
    code-tools: true
    css: styles.css
categories:
  - challenge_1
  - solution
---

Before starting, the necessary packages should be loaded:

```{r}
#| label: setup
#| warning: false
#| message: false

library(tidyverse)
library(igraph)

knitr::opts_chunk$set(echo = TRUE, warning=FALSE,
                      message=FALSE)
```

### Marriages in Game of Thrones

```{r}
got.marriages <- read_csv('_data/got/got_marriages.csv')

head(got.marriages)
```

The first two columns indicate the list of edges in the network. The network is symmetric and unweighted, but with 3 types of links (married, engaged, or affair):

```{r}
table(got.marriages$Type)
```

Finally, a graph can be created using the `igraph` package

```{r}
g.got.marriages <- graph_from_data_frame(got.marriages)
```

and plotted:

```{r}
plot(g.got.marriages, edge.arrow.size = 0)
```

We could also color the edges depending on the type of link by creating another column as a factor and changing its levels to color names:

```{r}
got.marriages$Type.Color <- as.factor(got.marriages$Type)

levels(got.marriages$Type.Color)

levels(got.marriages$Type.Color) <- c('gray', 'blue', 'green')

E(g.got.marriages)$color <- got.marriages$Type.Color

plot(g.got.marriages, edge.arrow.size = 0)
```

### Fish Encounters

`fish_encounters` dataset is directly loaded with `tidyverse` or `tidyr`:

```{r}
head(fish_encounters)
```

The graph can be created using:

```{r}
g.fish <- graph_from_data_frame(fish_encounters)

V(g.fish)$name
```

We can color the stations by selecting the nodes with names that cannot be converted into numbers:

```{r}
V(g.fish)$color[is.na(as.numeric(V(g.fish)$name))] <- 'lightblue'

plot(g.fish, edge.arrow.size = .5)
```

### Likes and Dislikes in Game of Thrones

```{r}
got.likes <- read.csv('_data/got/got_like_dislike.csv')

head(got.likes)
```

Although the data is given in the form of an adjacency matrix, the object is a data frame that needs to be coerced into a matrix. The network is asymmetric and weighted, where negative weights indicate dislike.

```{r}
g.likes <- graph_from_adjacency_matrix(as.matrix(got.likes[ , - c(1 : 3)]), mode = 'directed', weighted = TRUE)

kableExtra :: kable(as.matrix(got.likes[ , - c(1 : 3)]))
```

Characters in Game of Thrones seem to dislike each other much more than like each other...

```{r}
hist(E(g.likes)$weight)

E(g.likes)$color <- colorRampPalette(c("yellow", "blue"))(11)[E(g.likes)$weight + 6]

plot(g.likes, edge.arrow.size = 0.5)
```
