---
title: "Week 2 Challenge Instructions"
author: "Yakub Rabiutheen"
description: "Describing the Basic Structure of a Network"
date: "03/5/2023"
format:
  html:
    toc: true
    code-fold: true
    code-copy: true
    code-tools: true
# editor: visual
categories: 
  - challenge_2
  # - railroads
  # - faostat
  # - wildbirds
---

```{r}
library(readr)
library(igraph)
```

## Challenge Overview

For my Challenges , I will be doing the Game of Thrones Likes vs Dislike Dataset.


```{r}
got_like_dislike <- read_csv("_data/got/got_like_dislike.csv")
```
1. List and inspect List the objects to make sure the datafiles are working properly.

Below are the list of Objects in the Game of Thrones Like_Dislike Dataset
```{r}

```


2.Network Size What is the size of the network? You may use vcount and ecount.

First I will transform this into an igraph object

```{r}
got_like_dislike.ig <- graph_from_data_frame(got_like_dislike)
```

```{r}
vcount(got_like_dislike.ig)
```
```{r}
ecount(got_like_dislike.ig)
```



This Network has 11 vertices and 46 edges.

3.Network features Are these networks weighted, directed, and bipartite?


```{r}
is.bipartite(got_like_dislike.ig)
```


```{r}
is.directed(got_like_dislike.ig)
```


```{r}
is.weighted(got_like_dislike.ig)
```

This Network is a Directed, NonWeigthed and NonBipartite netowrk.


4.Network Attributes Now, using commands from either statnet or igraph, list the vertex and edge attributes.

Below are the Network Attributes
```{r}
igraph::vertex_attr_names(got_like_dislike.ig)
```
Here are the Names of Characters being used in Dislike vs Like
```{r}
igraph::edge_attr_names(got_like_dislike.ig)
```

Dyad Census
```{r}
igraph::dyad.census(got_like_dislike.ig)
```
It's interesting that  lot of Null values, which means there are Characters that no one Likes or Dislikes. Which would make sense as some characters such as Khal Dago don't have any interaction with most GOT characters.

Transtivity
```{r}
transitivity(got_like_dislike.ig)
```
```{r}
transitivity(got_like_dislike.ig, type = 'global')
```

A Global Transitvity of 0.2368 indicates that there are some local clusters or communities in the network, but they are not highly interconnected or cohesive.





```{r}
transitivity(got_like_dislike.ig, type = 'average') 
```
The  average transitivity value of 0.5867347 suggests that there may be certain nodes or subgroups of nodes in the network that have a much higher level of clustering or tendency to form triangle.

Now  I will look at Path Length , Number of Components and number of nodes in each connected component of the network.

```{r}
average.path.length(got_like_dislike.ig,directed = T)
igraph::components(got_like_dislike.ig)$no
igraph::components(got_like_dislike.ig)$csize
```
1.The average path length of 1.3 indicates  relatively small and tightly interconnected network.
2. The value of 1 indicates that there is only one connected component in the network, which means that all nodes in the network are connected to each other by paths.
3.There are 11 nodes in the largest connected component of the network. This indicates that the network is highly connected, and there are no isolated nodes or subgroups of nodes that are disconnected from the rest of the network.
